Dependently Typed Languages in Statix

### 2023-03-22 pres-evcs/presentation.tex
Work a bit on thesis
Update presentation

\begin{frame}[fragile]{Background: What are Dependent Types?}
\begin{itemize}
	\item Types may depend on values!
	\begin{exampleblock}{Example}
		\texttt{concat : (A: Set) -> (n m : Nat) -> Vec A n -> Vec A m
			\\ \hspace*{48pt} -> Vec A (n + m)}
	\end{exampleblock}
% - n is a RUNTIME value, it does not have to be known at compile time
% - n occurs insie a type `Vec`!
% - The type system enforces these lengths are always correct
% - Challenge: Type checking may require evaluating arbitrary terms to decide equality
	\item Proof assistants
	\item For example: Agda, Coq, Lean, ...
% Can be used to write proofs about code
\end{itemize}

@@

\begin{frame}[fragile]{Research Question}
\large{How suitable is Statix for defining a dependently-typed language?}
\begin{itemize}
	\item Defining system F was a challenge\footnote{Hendrik van Antwerpen et al. Scopes as types.}, will it even be possible?
	\item Will it be easier than doing it in Haskell?
\end{itemize}

@@

	\begin{exampleblock}{Example 1}
		\texttt{($\backslash$v: Type. v) T}
	\end{exampleblock}

	\begin{exampleblock}{Example 2}
		\texttt{let f = $\backslash$T: Type. $\backslash$x: T. x; \\
f (T: Type -> Type) ($\backslash$y: Type. y)
		}
	\end{exampleblock}
\end{frame}

\begin{frame}[fragile]{Type Checking}
	\begin{block}{Type checking relation}
		\texttt{typeOfExpr : scope * Expr -> Expr}
	\end{block}
	% Returns an Expr, not a Type
	% Scope -> Two different types of nodes
	\begin{block}{How do we use scopes?}
		A scope is used as a replacement for an environment and a context..\\
		One relation \texttt{name $\rightarrow$ NameEntry}, NameEntry is either:
		\begin{itemize}
			\item NType: Stores a type (Corresponds to a context)
			\item NSubst: Stores a substitution (Corresponds to an environment)
		\end{itemize}
	\end{block}
\end{frame}

@@

\begin{frame}[fragile]{Conclusions}
	Spoofax is a great tool for developing dependently typed languages!
	\begin{itemize}
		\item We can use scopes to represent environments and contexts
		\item Statix can still use improvements
	\end{itemize}

>>>>

\begin{frame}[fragile]{Background: What are Dependent Types?}
\begin{itemize}
	\item Types may depend on values!
	\begin{exampleblock}{Example}
		\texttt{concat : (A: Set) -> (n m : Nat) -> Vec A n -> Vec A m
			\\ \hspace*{48pt} -> Vec A (n + m)}
	\end{exampleblock}
% - n occurs insie a type `Vec`!
% - The type system enforces these lengths are always correct
% - Challenge: Type checking may require evaluating arbitrary terms to decide equality
	\item Proof assistants
	\item For example: Agda, Coq, Lean, ...
% Can be used to write proofs about code
\end{itemize}

@@

\begin{frame}[fragile]{Research Question}
\large{How suitable is Statix for defining a dependently-typed language?}
\begin{itemize}
	\item Will it be easier than doing it in Haskell?
	\item Defining system F was a challenge\footnote{Hendrik van Antwerpen et al. Scopes as types.}, will it even be possible?
\end{itemize}

@@

	\begin{exampleblock}{Example 1}
		\texttt{($\backslash$v: Type. v) Type}
	\end{exampleblock}

	\begin{exampleblock}{Example 2}
		\texttt{let f = $\backslash$T: Type. $\backslash$x: T. x; \\
f (T: Type -> Type) ($\backslash$y: Type. y)
		}
	\end{exampleblock}

	% most simple dependently typed language: lambda calculus
	% what other languages use as a core
\end{frame}

\begin{frame}[fragile]{Type Checking}
	\begin{block}{Type checking relation}
		\texttt{typeOfExpr : scope * Expr -> Expr}
	\end{block}
	% Returns an Expr, not a Type
	% Scope -> Two different types of nodes
	\begin{block}{How do we use scopes?}
		A scope is used as a replacement for an environment and a context.\\
		One edge \verb|p|. One relation \texttt{name $\rightarrow$ NameEntry}, NameEntry is either:
		\begin{itemize}
			\item NType: Stores a type (Corresponds to a context)
			\item NSubst: Stores a substitution (Corresponds to an environment)
		\end{itemize}
	\end{block}
\end{frame}

@@

\begin{frame}[fragile]{Conclusions}
	Spoofax is a great tool for developing dependently typed languages!
	\begin{itemize}
		\item We can use scopes to represent environments and contexts
		\item Statix can still use improvements
		% - Higher order 
	\end{itemize}
